The onset of hydrodynamically driven waving of marine vegetation in flow

### 2014-10-02 Paper/grass.tex
three figure per row
pulling

We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency, confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
We also demonstrate that our instability is different from the Kelvin-Helmholtz instability.

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.

\begin{figure*}
\begin{tabular}{cccc}
{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
\end{tabular}
\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
\label{K_Re_sigma_set3}
\end{figure*}

>>>>

We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency,
 confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
We demonstrate that flow instability mechanism leading to waving is different from the Kelvin-Helmholtz instability.

@@

We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.
\begin{figure*}
\begin{tabular}{cccc}
{\includegraphics[scale = .95]{Set4_dens28_imgsc}} &
%{\includegraphics[scale = 0.78]{Set4_dens30_imgsc}} &
{\includegraphics[scale = .95]{Set4_dens32_imgsc}} &
{\includegraphics[scale = .95]{Set4_dens34_imgsc}}\\
{\includegraphics[scale = .95]{Set5_dens38_imgsc}} &
%{\includegraphics[scale = 0.78]{Set5_dens40_imgsc}} &
{\includegraphics[scale = .95]{Set5_dens42_imgsc}} &
{\includegraphics[scale = .95]{Set5_dens46_imgsc}} \\
%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
\end{tabular}
\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
\label{K_Re_sigma_set3}
\end{figure*}
